---
title: "initial_exploration"
output: pdf_document
---

# Librairies and data import

We start by importing the Hadleyverse and importing our data.

```{r, message=FALSE, results = 'hide', warning=FALSE}
libs <- c('readr', 'lubridate', 'magrittr', 'tidyr', 'dplyr')
sapply(libs, require, character.only = TRUE)
train <- read_csv('~/Documents/kaggle_data/kaggle_rossmann/data/train.csv')
test <- read_csv('~/Documents/kaggle_data/kaggle_rossmann/data/test.csv')
stores <- read_csv('~/Documents/kaggle_data/kaggle_rossmann/data/store.csv')
```

We start by examining the train and stores data (the test data is structured as the train).

```{r}
head(train, 3)
head(stores, 3)
```

The train set contains information on **Sales**, which is the data to predict (the gross volume of cash made by a store during a day), while the stores data contains informations about the stores (distance to their competition for example).

We start by joining the datasets using the join method of dplyr.

```{r}
train <- inner_join(train, stores, by = 'Store')
test <- inner_join(test, stores, by = 'Store')
head(train, 3)
```

# NA filling

This sections explores basic strategies for NA filling. First we use the na_proportion function (defined in utilities) to show which columns have NAs and in which proportion.

```{r}
na_proportion <- function(dataset, digits = 2) {
  # returns the proportion of missing values
  # in each column of dataset, for which there is NAs
  na_col <- names(which(apply(dataset, 2, function(x) any(is.na(x))) == TRUE))
  if (length(na_col) >= 1) {
    sapply(select(dataset, one_of(na_col)), 
           function(x) round(sum(is.na(x)) / length(x), digits = digits))
  } else {
    cat("No missing values.")
  }
}
na_proportion(train)
```

## Competition Distance
We see a few missing values in the CompetitionDistance field, we can show explicitly those shops which lacks this information.

```{r}
filter(stores, is.na(CompetitionDistance)) %>% select(Store, CompetitionDistance)
```

It seems reasonable to imagine that those three stores do not have any competition? 

We make the disctuable assumption to fill those values by the maximum distance to the competition observed.

```{r}
max_comp_dist <- max(stores$CompetitionDistance, na.rm = TRUE)
train[is.na(train$CompetitionDistance), 'CompetitionDistance'] <- max_comp_dist
test[is.na(test$CompetitionDistance), 'CompetitionDistance'] <- max_comp_dist
```

## Some shop in the test set are closed

It should be noticed that some Sales to be predicted are for closed shops.

```{r}
test %>% filter(Open == 0) %>% head(3)
```

We need to guarantee that for those days our prediction are 0. To do so we define a function in utilities which should **always** be applied to any prediction vector.

```{r}
set_pred_closed <- function(pred_vec) {
  # this function set to zero the Sales predictions
  # anytime the shop is closed in the test set
  pred_vec[which(test$Open == 0)] <- 0
}
```


## Some shop have missing Open value

```{r}
test %>% filter(is.na(Open)) %>% head(3)
```

To see what the filling should be we need to inspect the corresponding dates. We can notice already that the Promo takes the value 1 for some of those days: so even the shop is open and there was a bug, or the shop was in promo and closed !

Let us do saome date aprsing (using lubridate and the fact that this is a german retailer).

```{r}
train$Date <- ymd(train$Date, tz = "Europe/Berlin")
test$Date <- ymd(test$Date, tz = "Europe/Berlin")
```
 Now we can explore the days where we see that anomaly.
 
 
```{r}
test_open_anomaly <- filter(test, is.na(Open))
wday(test_open_anomaly$Date, label = TRUE)
```

No sunday ! When is this shop usually closed ?

```{r}
store_622 <- filter(test, Store == 622)
store_622_closed <- filter(store_622, Open == 0, StateHoliday == 0)
unique(wday(store_622_closed$Date, label = TRUE))
```

It is only closed on Sunday ! Then we decide to fill those values with 1's (i.e. the shop was probably open).

```{r}
test[is.na(test$Open), 'Open'] <- 1
```

## The state holiday problem

StateHoliday never takes the value 1!

```{r}
sum(train$StateHoliday == 0, na.rm = TRUE)
sum(train$StateHoliday == 1, na.rm = TRUE) # jamais 1!
sum(test$StateHoliday == 0, na.rm = TRUE)
sum(test$StateHoliday == 1, na.rm = TRUE) # jamais 1!
```

Let us look at some of those rows where StateHoliday is missing.

```{r}
filter(train, is.na(StateHoliday)) %>% select(Date)
```

Those dates are probably holidays, so we decide to fill by 1 all those StateHoliday missing values.

```{r}
train[is.na(train$StateHoliday), 'StateHoliday'] <- 1
test[is.na(test$StateHoliday), 'StateHoliday'] <- 1
```


